---
title: "Homework 2"
format: pdf
  mainfont: Garamound
editor: visual
---

## 1 Child Mortality in Cote d'Ivoire

```{r}
## Libraries
library(haven)
library(tidyverse)
library(knitr)
library(kableExtra)
library(tidyr)
library(Amelia)
library(sandwich)
library(lmtest)
library(miceadds)
library(huxtable)
library(matrixStats)
library(float)

```

```{r}
mortality <- read_dta("civ88_mort.dta")
nutri <- read_dta("civ88-08_nutri.dta")
sum(is.na(mortality)) #=0
sum(is.na(nutri)) #=12474

missing_df<-as.data.frame((missingvalues))
missing_df%>%
  filter(missing_df!=0)%>%
  kable(caption = "Number of missing values per variable",
        col.names=c("Variables", "Missing values"))%>%
  kable_classic()

#Are the missing values random?
dummies<-nutri%>%
  select(height, weightkg)%>%
  mutate(dummyheight=ifelse(is.na(height),1,0),
         dummyweightkg=ifelse(is.na(weightkg),1,0))

heightmv<-lm(nutri$conspc88~dummies$dummyheight)
weightmv<-lm(nutri$conspc88~dummies$dummyweightkg)

indepheight<-huxreg("Regression of height on variables of the dataset"= 
                      lm(nutri$height ~ nutri$hhid + nutri$hhsize + nutri$north + nutri$age + nutri$sexe + nutri$education + nutri$medheight + nutri$conspc88, data=nutri))

indepheight

indepweight<-huxreg("Regression of weight on variables of the dataset"= 
                      lm(nutri$weightkg ~ nutri$hhid + nutri$hhsize + nutri$north + nutri$age + nutri$sexe + nutri$education + nutri$medheight + nutri$conspc88, data=nutri))

indepweight

newnutri<-nutri%>%
  mutate(height=ifelse(is.na(height)==T, 0.952*medheight, height), 
         weightkg=ifelse(is.na(weightkg)==T, 50.467-0.330*hhsize+0.456*age+10.388*education-0.388*medheight, weightkg))
```

### 1.1. Sampling

```{r}
mortality <- mortality %>% mutate(weight = (pweight*n_wom))

```

### 1.2. Measurement of under 5 mortality

Either take the population of children born 0 to 4 years ago and divide the number of deaths by the number of children

**b.1.) First Method**

```{r}
## make prettier tables
## Compute the average
# Boys
summary(lm(data = mortality, formula = dboys_l5/boys_l5 ~ 1, weights = weight))
# Girls
summary(lm(data = mortality, formula = dgirls_l5/girls_l5 ~ 1, weights = weight))

```

Or take the population of children born 5 to 9 years ago and divide the number of deaths between 0 and 4 years of age (i.e. before reaching age 5) by the number of children.

**b.2.) Second Method**

```{r}
## Numer of children born 5 to 9 years ago
mortality <- mortality %>%  mutate(boys_5_9= boys - boys_l5)
mortality <- mortality %>% mutate(girls_5_9 = girls - girls_l5)

## Number of children dead before age of five born between 0 and 5 years ago
mortality <- mortality <- mortality %>%  mutate(dboyl9 = dboys_u5 - dboys_l5 )
mortality <- mortality <- mortality %>%  mutate(dgirlsl9 = dgirls_u5 - dgirls_l5 )

## Weighted average
# Boys = 0.120314
summary(lm(data = mortality, formula = dboyl9/boys_5_9 ~ 1, weights = weight, na.rm=T))
# Girls = 0.116958
summary(lm(data = mortality, formula = dgirlsl9/girls_5_9 ~ 1, weights = weight, na.rm=T))

##Tables (kind of obvious that it was done w Chatgpt, feel free to change if you have better ideas!
extract <- function(x) {
  coef <- coef(summary(x))[1, 1]
  se <- coef(summary(x))[1, 2]
  return(c(mean = coef, se = se))
}

m1_gstats<-extract(m1_g)
m1_bstats<-extract(m1_b)
m2_gstats<-extract(m2_g)
m2_bstats<-extract(m2_b)

m1_g <- paste0(round(m1_gstats["mean"], 3), " (", round(m1_gstats["se"], 3), ")")
m1_b <- paste0(round(m1_bstats["mean"], 3), " (", round(m1_bstats["se"], 3), ")")
m2_g <- paste0(round(m2_gstats["mean"], 3), " (", round(m2_gstats["se"], 3), ")")
m2_b <- paste0(round(m2_bstats["mean"], 3), " (", round(m2_bstats["se"], 3), ")")

child_mort<-data.frame(
  "Method 1: Based on children born 0 to 4 years ago" = c(m1_g, m1_b), 
  "Method 2: Based on children born 5 to 9 years ago" = c(m2_g, 
                                                          m2_b),
  row.names = c("Girls", "Boys"))

kable(child_mort, caption = "Under-5 mortality rates in Côte d'Ivoire")

### Compare SE
```

## 2. Health status, nutritional status and wealth in Cote d'Ivoire

```{r}
civ88_08_nutri <- read_dta("civ88-08_nutri.dta")
nutri <- civ88_08_nutri
kable(summary(nutri$year))
```

### 2.1. Units

In 2008, the weight could be measured in either kilograms or hectograms. However, the unit was not systematically reported. It was therefore assumed that all the weights were reported in kilograms.

**a) Sample mean**

Derive an expression for the sample mean weight in this context, and show that, if we knew just one parameter, we could obtain a non-biased estimate of the mean weight despite the hectogram issue. (.5 point)

$\bar x = \frac{\sum x_\text{i hect}*0.01 + \sum x_\text{i kg}} {\sum n}$

```{r}
par(mfrow=c(1,3))
weight2008<- nutri %>% filter(year==2008) %>% select(weightkg)
 ggplot(weight2008, aes(weightkg)) + geom_histogram(col="darkblue", fill="blue") + theme_classic()
 
 weight1988<- nutri %>% filter(year==1988) %>% select(weightkg)
 ggplot(weight1988, aes(weightkg)) + geom_histogram(col="darkblue", fill="darkgreen") + theme_classic()
 
 weight1993<- nutri %>% filter(year==1993) %>% select(weightkg)
ggplot(weight1993, aes(weightkg)) + geom_histogram(col="darkblue", fill="purple") + theme_classic()
```

nutri %\>% filter(year==1993) %\>% select(weightkg)

ggplot(weight1993, aes(weightkg)) + geom_histogram(aes(y=..density..), \# Histogram with density instead of count on y-axis

binwidth=.5,

colour="black", fill="white") +

geom_density(alpha=.2, fill="#FF6666")

```{r}
nutri %>% filter(year==2008) %>% select(weightkg)
 ggplot(weight2008, aes(weightkg)) +  geom_histogram(aes(y=..density..),      # Histogram with density instead of count on y-axis
                   binwidth=.5,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666") +
    geom_vline(aes(xintercept=mean(weightkg, na.rm=T)),   # Ignore NA values for mean
               color="red", linetype="dashed", size=1)
 
nutri %>% filter(year==1988) %>% select(weightkg)
 ggplot(weight1988, aes(weightkg)) + ggeom_histogram(aes(y=..density..),      # Histogram with density instead of count on y-axis
                   binwidth=.5,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")
 
nutri %>% filter(year==1993) %>% select(weightkg)
 ggplot(weight1988, aes(weightkg)) + ggeom_histogram(aes(y=..density..),      # Histogram with density instead of count on y-axis
                   binwidth=.5,
                   colour="black", fill="white") +
    geom_density(alpha=.2, fill="#FF6666")
```

I think we can make them even prettier <http://www.cookbook-r.com/Graphs/Plotting_distributions_(ggplot2)/>

**b) Abnormal values**

[Weights for age from the WHO website](https://www.who.int/tools/child-growth-standards/standards/weight-for-age)

```{r}
ages<-c("0 to 1 y.o.", "1 to 2 y.o.", "2 to 3 y.o.", "3 to 5 y.o.")
threshw<-c("up to 12 kg", "up to 15 kg", "up to 18 kg", "up to 24.5 kg")
thresholds<-data.frame("Age" = ages, "Normal weights" = threshw)
kable(thresholds, caption="Expected weights of children from birth to 5 years
      old in the world")


nutri_kg <- nutri %>%
  mutate(weightkg = case_when(
    year == "2008" & age <= 12 & weightkg > 12 ~ weightkg * 0.1,
    year == "2008" & age <= 12 ~ weightkg,
    year == "2008" & age > 12 & age <= 24 & weightkg > 15 ~ weightkg * 0.1,
    year == "2008" & age > 12 & age <= 24 ~ weightkg,
    year == "2008" & age > 24 & age <= 36 & weightkg > 18 ~ weightkg * 0.1,
    year == "2008" & age > 24 & age <= 36 ~ weightkg,
    year == "2008" & age > 36 & weightkg > 24.5 ~ weightkg * 0.1,
    year == "2008" & age > 36 ~ weightkg,
    TRUE ~ weightkg
  ))
```

```{r}


kable(nutri %>% summary(age))
nutri %>% max(age) ##= 60 months = 5 years
## weights higher than 30kg -> switch to kg
nutri <- nutri %>% mutate(weightkg= ifelse(weightkg > 30, weightkg*0.1, weightkg*1))

## weights lower than 1 kg and higher than 30 kg return NA
nutri <- nutri %>% mutate(weightkg= ifelse(weightkg > 30 | weightkg<1, NA, weightkg*1))
```

**Graphical representation**

```{r}
par(mfrow=c(1,3))
weight2008<- nutri %>% filter(year==2008) %>% select(weightkg)
 ggplot(weight2008, aes(weightkg)) + geom_histogram(col="darkblue", fill="blue") + theme_classic()
 
 weight1988<- nutri %>% filter(year==1988) %>% select(weightkg)
 ggplot(weight1988, aes(weightkg)) + geom_histogram(col="darkblue", fill="darkgreen") + theme_classic()
 
 weight1993<- nutri %>% filter(year==1993) %>% select(weightkg)
ggplot(weight1993, aes(weightkg)) + geom_histogram(col="darkblue", fill="purple") + theme_classic()
```

WE could do this for age groups since some weights do make sense in absolute terms but not when compared with the respective age.

### 2.2. Z Scores

Using the variables medheight and stdheight, compute each child's z-score.

$z = \frac {height - WHOMedian (sex, age)} {WHOstddev (sex,age)}$

```{r}
nutri2 <- nutri %>% mutate(zscore= (height- medheight)/stdheight)
summary(nutri2$zscore)
```

```{r}
ggplot(nutri2, aes(x = zscore)) +
  geom_histogram(binwidth = 0.5, fill = "pink", color = "black") +
  ggtitle("Histogram zscore") +
  xlab("zscore") +
  xlim(-5, 5)  +
    geom_density(alpha=.2, fill="#FF6666") + theme_bw()

```

### 3. Height Stature over time and across regions

*(a) Estimate the stunting rate for 2--4 years old children over time in the North, in the South, and at the national level. Decide what are outliers, and show results both including and excluding outliers. Comment your results.*

```{r}
## stunting rate
nutri2 <- nutri2 %>% mutate(stunned = ifelse(zscore <= -2, 1, 0))
##!! children 2-4 south, north national level

## outliers
nutri3 <- nutri2  %>% mutate(zscore = ifelse(zscore <= -3| zscore > 3, NA, zscore
)) %>% mutate(stunned = ifelse(zscore <= -2, 1, 0))
summary(nutri3$zscore)

## recalculate rates
nutri2 %>% filter(year==2008)%>%   filter(!is.na(stunned)) %>% 
  filter(age > 24 & age < 48)  %>%  summarize(obs=n(),prop=sum(stunned==1)/n())
  
nutri2 %>% filter(year==1988)%>%   filter(!is.na(stunned)) %>% 
  filter(age > 24 & age < 48) %>%  summarize(obs=n(),prop=sum(stunned==1)/n())

nutri2 %>% filter(year==1993)%>%
  filter(age > 24 & age < 48) %>%   filter(!is.na(stunned)) %>%  summarize(obs=n(),prop=sum(stunned==1)/n())

## or
prop <- nutri2 %>% sum(nutri2$stunned==1)/sum(nutri2)
nutri2 %>%  filter(!is.na(stunned)) %>% (lm(data=nutri2, formula= prop ~ 1))
## does not work
```

*(b) Estimate the stunting rate for 2--4 years old children over time for boys and for girls at the national
level. Decide what are outliers, and show results both including and excluding outliers. Comment your results.*

**4. Height Stature and poverty (2 points)**
Here, one would like to study how the gradient of the number of stunted children aged 2--4 years, and of household consumption per capita, as well as the relationship between the two, have changed over time.
Find a graphical or regression framework to represent this (many options are possible). Comment on your results and their limits.

-   regression between stunted children, consumption per capita in different years

-   graphs on evolution of both

-   plotting regression

**5. Immigration**

Crime rate and immigration?
See if current legal immigration to European countries is coming more from previous colonies (too many omitted variables). 
Link between immigration and far-right prevalence. 
